---
title: "Analysis of the TME in the cohort of bc-meta"
output: html_notebook
params:
    seurat_object_bpcells: "seu_lvl2_merged.Rds"
---

> Packages

```{r, setup, message=FALSE}

# # Install packages #
# install.packages('Seurat')
# 
# setRepositories(ind = 1:3, addURLs = c('https://satijalab.r-universe.dev', 'https://bnprks.r-universe.dev/'))
# install.packages(c("BPCells", "presto", "glmGamPoi"))
# 
# install.packages('Signac')
# remotes::install_github("satijalab/seurat-data", quiet = TRUE)
# remotes::install_github("satijalab/azimuth", quiet = TRUE)
# remotes::install_github("satijalab/seurat-wrappers", quiet = TRUE)

# load packages #

library('dplyr')
library('BPCells')
library('Seurat')
library('SeuratObject')
library('SeuratDisk')
#BiocManager::install("rhdf5")
#library('Azimuth')
library('SeuratData')
library('ggplot2')

options(Seurat.object.assay.version = "v5")
options(future.globals.maxSize = 891289600) # (850 * 1024^2)
knitr::opts_knit$set(root.dir = './data/merged/')

# tcca palette
tumor_sites_colors <- c("brain" = "#CA9A8C",
                        "adrenal_gland" = "#FF9A4A",
                        "breast" = "#FF8FAB",
                        "skin" = "#5E2D2C",
                        "lung" = "#A2F0EF",
                        "soft_tissue" = "#FB6467",
                        "esophagus" = "#5780FE", 
                        "bladder" = "#58B368",
                        "lymph_node" = "#B6F884",
                        "liver" = "#309975",
                        "pancreas" = "#B47EB3",
                        "ovary" = "#9C0D38",
                        "prostate" = "#005D95",
                        "other" = "#BBB9B7")

plot_umap <- function(x, labels, raw_mat_df, feature,
                      colors=c("#ff7f00", "#e377c2", "#17becf"),
                      pad=0.1, cex=0.6, pch=19, legend.suffix="",
                      cex.main=1, cex.legend=0.85) {
  
  layout <- x
  if (is(x, "umap")) {
    layout <- x$layout %>% as.data.frame() %>% tibble::rownames_to_column(var = "sample")
    layout_df <- dplyr::inner_join(layout, labels, by = "sample")
    layout_df <- dplyr::inner_join(layout_df, raw_mat_df, by = c("sample", "consensus_hclust"))
    layout_df <- layout_df %>% dplyr::filter(TME_feature == "Stromal")
  }
  # 
  # xylim <- range(layout)
  # xylim <- xylim + ((xylim[2]-xylim[1])*pad)*c(-0.5, 0.5)

  plot <- layout_df %>% 
    ggplot(aes(x=V1, y=V2, colour = consensus_hclust)) +
    geom_point() +
    scale_color_manual(values = colors_clustering[1:length(unique(layout_df$consensus_hclust))]) +
    theme_classic() +
    theme(axis.text = element_text(size = 12, color = "black")) +
    labs(x = "UMAP 1", y = "UMAP 2")
  
  if(!is.null(feature)){
    plot <- layout_df %>% 
      ggplot(aes(x=V1, y=V2, colour = frac)) +
      geom_point() +
      scale_color_continuous(type = "viridis") +
      theme_classic() +
      theme(axis.text = element_text(size = 12, color = "black")) +
      labs(x = "UMAP 1", y = "UMAP 2")
  }
      
    # par(mar=c(1,1,1,1), ps=10)
    # plot(xylim, xylim, type="n", axes=F, frame=F)
    # rect(xylim[1], xylim[1], xylim[2], xylim[2], border="black", lwd=1)
  
  # points(layout[,1], layout[,2], col=colors[as.integer(labels)],
  #        cex=cex, pch=pch)
  # mtext(side=1, "UMAP 1", cex=cex.main)
  # mtext(side=2, "UMAP 2", cex=cex.main)
  
  # labels.u <- unique(labels)
  # legend.pos <- "topleft"
  # legend.text <- as.character(labels.u)
  # if (add) {
  #   legend.pos <- "bottomleft"
  #   legend.text <- paste(as.character(labels.u), legend.suffix)
  # }
  # 
  # legend(legend.pos, legend=legend.text, inset=0.03,
  #        col=colors[as.integer(labels.u)],
  #        bty="n", pch=pch, cex=cex.legend)
  
  return(plot)
}

```

> Read data

```{r, include=FALSE}

## read merged bc-meta seurat object, 
merged_obj <- readRDS(params$seurat_object_bpcells)
lvl2_metadata <- merged_obj@meta.data
SeuratObject::Version(merged_obj)

# .......................................................... #
# Create new seurat object pointing to local directory
# .......................................................... #
# # open matrix from directory
# mat <- open_matrix_dir(dir = "./data/merged_bc_meta_counts")
# 
# # overwrite current directory to previos directory
# merged_bc_metanalysis_full[["RNA"]]$counts <- mat
# 
# # save new seurat object pointing to currect directory
# SeuratObject::saveRDS(
#   object = merged_bc_metanalysis_full,
#   file = "merged_bc_metanalysis_full.Rds",
#   destdir = "./data"
# )

```

> Cell-type harmonization (low_res)

- Mapping QC based on the provided mapping and prediction scores.

1. The mapping scores are used to assess how whether the query cell type is well represented in the reference.

2. The prediction scores are used to assess how confident are we about the annotation.

* I would say that is more relevant to have a good reference rather than a closest cell type prediction. 

- Cell types are harmonized regarding two different degrees of resolution.

1. Low resolution: adiposeref_predicted.celltype.l1.score, tonsilref_predicted.celltype.l1.score, lungref_predicted.ann_level_4.score, pbmcref_predicted.celltype.l1.score, humancortexref_predicted.subclass.score, bonemarrowref_predicted.celltype.l2.score

2. High resolution:

```{r}

#### Fix "duplicate"d T19 sample (adrenalnb_rui_chong & pdac_junya_peng)
### Keep just pdac_junya_peng, as the other one cannot be found in cell-level annotations
lvl2_metadata[
    lvl2_metadata$sample == "T19" &
    lvl2_metadata$study == "adrenalnb_rui_chong",
    "sample"
    ] <- "T19_1"

merged_obj@meta.data <- lvl2_metadata

#### Remove ambiguous BRCA samples (tumor_type == "BRCA" & tumor_subtype == "predicted_tumor, ~ 20 samples)
merged_obj <- subset(merged_obj, subset = tumor_type == 'BRCA' & tumor_subtype == 'predicted_tumour', invert = TRUE)

# Automatic cell-type annotation using Azimuth ---------------------------------

## keep non-malignant cells (673 samples)
merged_obj_M <- subset(merged_obj, malignancy == TRUE)
bcmeta_annot_M <- merged_obj_M@meta.data
merged_obj_nM <- subset(merged_obj, malignancy == FALSE)
bcmeta_annot_nM <- merged_obj_nM@meta.data

## load azimuth references using SeuratData
available_data <- SeuratData::AvailableData()
available_data[grep("Azimuth", available_data[, 3]), 1:3]

name_refs <- c("pbmcref", "lungref", "pancreasref", "tonsilref", "adiposeref", "humancortexref", "bonemarrowref")
# "kidneyref" not used (open issue in github, also original study normalizes data in a different way to seurat standards)

## collect annotation results from cluster
metadata_TME <- do.call(cbind, lapply(name_refs, function(X){
  tmp_obj <- readRDS(file = paste0("../azimuth2/azimuth_annotation_TME_", X,".rds"))
  tmp_metadata <- tmp_obj@meta.data
  tmp_metadata <- tmp_metadata[, grep("predicted|mapping", colnames(tmp_metadata))]
  colnames(tmp_metadata) <- paste0(X, "_", colnames(tmp_metadata))
  return(tmp_metadata)
}))

metadata_TME <- cbind(bcmeta_annot_nM, metadata_TME)

## Note that level 4 in lung, display "none" in some cases. We keep finest level to solve this.

## Remove annotations that are too detailed
metadata_TME <- metadata_TME %>%
  select(!c(pbmcref_predicted.celltype.l3, pbmcref_predicted.celltype.l3.score, # pbmcref
            humancortexref_predicted.class, humancortexref_predicted.class.score, # humancortexref
            humancortexref_predicted.cluster, humancortexref_predicted.cluster.score, # humancortexref
            humancortexref_predicted.cross_species_cluster, humancortexref_predicted.cross_species_cluster.score, # humancortexref
            #pancreasref_predicted.annotation.l1, pancreasref_predicted.annotation.l1.score, pancreasref_mapping.score, # pancreasref
            lungref_predicted.ann_level_1, lungref_predicted.ann_level_1.score, # lungref
            lungref_predicted.ann_level_2, lungref_predicted.ann_level_2.score, # lungref
            lungref_predicted.ann_level_5, lungref_predicted.ann_level_5.score, # lungref
            #lungref_predicted.ann_finest_level, lungref_predicted.ann_finest_level.score, # lungref
            tonsilref_predicted.celltype.l2, tonsilref_predicted.celltype.l2.score, # tonsilref,
            adiposeref_predicted.celltype.l2, adiposeref_predicted.celltype.l2.score, # adiposeref
         ))

# Quality control metrics -----------------------------------------------------

### MAPPING: reflects confidence that this cell is well represented by the reference ####
tmp <- metadata_TME %>%
  dplyr::mutate(tumor_site = ifelse(metadata_TME$refined_tumor_site %in% names(tumor_sites_colors),
                                    refined_tumor_site, "other")) %>%
  dplyr::rename("adipose" =  "adiposeref_mapping.score",
                "tonsil" =  "tonsilref_mapping.score", 
                "lung" =  "lungref_mapping.score",
                "pbmc" =  "pbmcref_mapping.score",
                "humancortex" =  "humancortexref_mapping.score",
                "bonemarrow" =  "bonemarrowref_mapping.score",
                "pancreas" = "pancreasref_mapping.score") %>%
  dplyr::select(tumor_site, pbmc, lung, tonsil, adipose, humancortex, bonemarrow, pancreas) %>%
  tidyr::pivot_longer(cols = !tumor_site, names_to = "reference", values_to = "mapping_score")
  
tmp %>%
  ggplot(aes(x=reference, y=mapping_score, fill=tumor_site)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = tumor_sites_colors) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.y = element_text(size=10, color = "black"),
        axis.text.x = element_text(size=10, color = "black")) +
  labs(x = "Azimuth reference", y = "Mapping score", fill = "Tumor site")

ggsave(filename = "../../results/qc_mapping_azimuth2.png", width = 12, height = 8)

### PREDICTION: reflect the confidence associated with each annotation ####

tmp <- metadata_TME %>%
  dplyr::mutate(tumor_site = ifelse(metadata_TME$refined_tumor_site %in% names(tumor_sites_colors),
                                    refined_tumor_site, "other")) %>%
  dplyr::rename("adipose" =  "adiposeref_predicted.celltype.l1.score",
                "tonsil" =  "tonsilref_predicted.celltype.l1.score", 
                "lung" =  "lungref_predicted.ann_level_4.score",
                "lung_finest" = "lungref_predicted.ann_finest_level.score",
                "pbmc" =  "pbmcref_predicted.celltype.l1.score",
                "humancortex" =  "humancortexref_predicted.subclass.score",
                "bonemarrow" =  "bonemarrowref_predicted.celltype.l2.score",
                "pancreas" = "pancreasref_predicted.annotation.l1.score") %>%
  dplyr::select(tumor_site, pbmc, lung, tonsil, adipose, humancortex, bonemarrow, pancreas) %>%
  tidyr::pivot_longer(cols = !tumor_site, names_to = "reference", values_to = "prediction_score")
  
tmp %>%
  ggplot(aes(x=reference, y=prediction_score, fill=tumor_site)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = tumor_sites_colors) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.y = element_text(size=10, color = "black"),
        axis.text.x = element_text(size=10, color = "black")) +
  labs(x = "Azimuth reference", y = "Prediction score (Low resolution)", fill = "Tumor site")

ggsave(filename = "../../results/qc_prediction_low_res_azimuth2.png", width = 12, height = 8)

# Evaluate confidence of cell-type annotations ---------------------------------

## Mapping scores
mapping_df <- metadata_TME %>% 
  select(cell, sample, tumor_site, 
         pbmcref_mapping.score, 
         lungref_mapping.score,
         tonsilref_mapping.score, 
         adiposeref_mapping.score, 
         humancortexref_mapping.score, 
         bonemarrowref_mapping.score,
         pancreasref_mapping.score) %>%
  dplyr::rename("adipose" =  "adiposeref_mapping.score",
                "tonsil" =  "tonsilref_mapping.score", 
                "lung" =  "lungref_mapping.score",
                "pbmc" =  "pbmcref_mapping.score",
                "humancortex" =  "humancortexref_mapping.score",
                "bonemarrow" =  "bonemarrowref_mapping.score",
                "pancreas" = "pancreasref_mapping.score") %>%
  tidyr::pivot_longer(cols = !c(cell, sample, tumor_site),
                      values_to = "mapping_score", 
                      names_to = "reference") %>%
  dplyr::mutate(tumor_site = ifelse(tumor_site %in% names(tumor_sites_colors), tumor_site, "other")) %>%
  dplyr::mutate(tumor_site = factor(tumor_site, levels = names(tumor_sites_colors)))

## Distribution
mapping_df %>%
  ggplot(aes(x=mapping_score, fill = tumor_site)) +
  geom_histogram(bins = 100) +
  scale_fill_manual(values = tumor_sites_colors) +
  facet_wrap(. ~ reference, scales = "free")

## Prediction scores
prediction_scores_df <- metadata_TME %>% 
  select(cell, sample, tumor_site, 
         pbmcref_predicted.celltype.l1.score, 
         lungref_predicted.ann_finest_level.score,
         tonsilref_predicted.celltype.l1.score, 
         adiposeref_predicted.celltype.l1.score, 
         humancortexref_predicted.subclass.score, 
         bonemarrowref_predicted.celltype.l2.score,
         pancreasref_predicted.annotation.l1.score) %>%
  dplyr::rename("adipose" =  "adiposeref_predicted.celltype.l1.score",
                "tonsil" =  "tonsilref_predicted.celltype.l1.score", 
                "lung" =  "lungref_predicted.ann_finest_level.score",
                "pbmc" =  "pbmcref_predicted.celltype.l1.score",
                "humancortex" =  "humancortexref_predicted.subclass.score",
                "bonemarrow" =  "bonemarrowref_predicted.celltype.l2.score",
                "pancreas" = "pancreasref_predicted.annotation.l1.score") %>%
  tidyr::pivot_longer(cols = !c(cell, sample, tumor_site),
                      values_to = "prediction_score", 
                      names_to = "reference") %>%
  dplyr::mutate(tumor_site = ifelse(tumor_site %in% names(tumor_sites_colors), tumor_site, "other")) %>%
  dplyr::mutate(tumor_site = factor(tumor_site, levels = names(tumor_sites_colors)))

## Predicted cell-type labels
prediction_labels_df <- metadata_TME %>% 
  select(cell, sample, cell_type, 
         pbmcref_predicted.celltype.l2, 
         lungref_predicted.ann_finest_level,
         tonsilref_predicted.celltype.l1, 
         adiposeref_predicted.celltype.l1, 
         humancortexref_predicted.subclass, 
         bonemarrowref_predicted.celltype.l2,
         pancreasref_predicted.annotation.l1) %>%
  dplyr::rename("adipose" =  "adiposeref_predicted.celltype.l1",
                "tonsil" =  "tonsilref_predicted.celltype.l1", 
                "lung" =  "lungref_predicted.ann_finest_level",
                "pbmc" =  "pbmcref_predicted.celltype.l2",
                "humancortex" = "humancortexref_predicted.subclass",
                "bonemarrow" = "bonemarrowref_predicted.celltype.l2",
                "pancreas" = "pancreasref_predicted.annotation.l1",
                "author_cell_type" = "cell_type") %>%
  tidyr::pivot_longer(cols = !c(cell, sample, author_cell_type),
                      values_to = "predicted_cell_type", 
                      names_to = "reference")

## Put together mapping and prediction information
mapping_df$prediction_score <- prediction_scores_df$prediction_score
mapping_df$author_cell_type <- prediction_labels_df$author_cell_type
mapping_df$predicted_cell_type <- prediction_labels_df$predicted_cell_type

## Add confidence variable for azimuth annotation: "annotated_with_azimuth" (threshold 0.7 based on visualization)
mapping_df <- mapping_df %>%
  dplyr::mutate(azimuth_annotation = ifelse(mapping_score >= 0.7, TRUE, FALSE))

### Save object
saveRDS(mapping_df, file = "../azimuth2/azimuth_analysis.rds")

### Add information to metadta object
metadata_TME_v2 <- metadata_TME[, 1:29]

### Save full metadata object
saveRDS(metadata_TME_v2, file = "../azimuth2/refined_metadata_TME.rds")

## Cell-type harmonization: 1st step manual curation ---------------------------

## Correspondence author vs azimuth mapping ###
author_vs_azimuth <- mapping_df %>% 
  group_by(author_cell_type, predicted_cell_type) %>% 
  summarise()

### Author's ### (Check .csv files afterwards to remove inconsistencies)

# # ** B cells **
# B <- c("B cell", "B.cell", "B-cell", "B_cell", "B cells", "B-cells", "B_cells", "B.cells", "B.cell", "B.Cell", "B")
# B_idx <- unlist(sapply(B, function(X) grep(X, author_vs_azimuth$author_cell_type)))
# B_ids <- unique(author_vs_azimuth$author_cell_type[B_idx])
# data.table::fwrite(as.data.frame(B_ids), file = "../azimuth/B_author_celltypes.csv", row.names = FALSE)
# 
# # ** T cells **
# Tcell <- c("CD4 T", "CD8 T", "other T", "Treg", "CD4", "CD8", "T_cell", "T-cell", "T-Reg", 
#            "Lymphocyte", "T cell", "T-Helper", "T.cell", "MAIT", "Tcell_Prolif","Tprolif", "TCells", "Tfh",
#            "T/NK", "T and NK", "NKT", "NK_T_cells", "T_NK", "T:")
# T_idx <- unlist(sapply(Tcell, function(X) grep(X, author_vs_azimuth$author_cell_type)))
# T_ids <- unique(author_vs_azimuth$author_cell_type[T_idx])
# data.table::fwrite(as.data.frame(T_ids), file = "../azimuth/Tcell_author_celltypes.csv", row.names = FALSE)
# 
# # ** NK cells ** (T and NK considered before as T cell)
# NK <- c("NK")
# NK_idx <- unlist(sapply(NK, function(X) grep(X, author_vs_azimuth$author_cell_type)))
# NK_ids <- unique(author_vs_azimuth$author_cell_type[NK_idx])
# data.table::fwrite(as.data.frame(NK_ids), file = "../azimuth/NK_author_celltypes.csv", row.names = FALSE)
# 
# # ** Plasma cells ** (B and Plasma considered as B cell before, Plasmablast considered separately)
# Plasma <- c("Plasma", "Plasmablast", "Plasmocyte", "plasmocyte")
# Plasma_idx <- unlist(sapply(Plasma, function(X) grep(X, author_vs_azimuth$author_cell_type)))
# Plasma_ids <- unique(author_vs_azimuth$author_cell_type[Plasma_idx])
# data.table::fwrite(as.data.frame(Plasma_ids), file = "../azimuth/Plasma_author_celltypes.csv", row.names = FALSE)
# 
# # ** Endothelial cells ** 
# Endo <- c("Endo", "EC", "EPC")
# Endo_idx <- unlist(sapply(Endo, function(X) grep(X, author_vs_azimuth$author_cell_type)))
# Endo_ids <- unique(author_vs_azimuth$author_cell_type[Endo_idx])
# data.table::fwrite(as.data.frame(Endo_ids), file = "../azimuth/Endothelial_author_celltypes.csv", row.names = FALSE)
# 
# # ** CAFs ** 
# CAF <- c("stromal", "fibro", "CAF", "Fibroblast", "fibroblast", "FB", "Fibro")
# CAF_idx <- unlist(sapply(CAF, function(X) grep(X, author_vs_azimuth$author_cell_type)))
# CAF_ids <- unique(author_vs_azimuth$author_cell_type[CAF_idx])
# data.table::fwrite(as.data.frame(CAF_ids), file = "../azimuth/CAFs_author_celltypes.csv", row.names = FALSE)
# 
# # ** DC ** 
# DC <- c("Dendritic", "DC")
# DC_idx <- unlist(sapply(DC, function(X) grep(X, author_vs_azimuth$author_cell_type)))
# DC_ids <- unique(author_vs_azimuth$author_cell_type[DC_idx])
# data.table::fwrite(as.data.frame(DC_ids), file = "../azimuth/DC_author_celltypes.csv", row.names = FALSE)
# 
# # ** Macrophages ** 
# M <- c("Mφ", "Macrophage", "macrophage", "mac", "TAM", "Macro", "Mac")
# M_idx <- unlist(sapply(M, function(X) grep(X, author_vs_azimuth$author_cell_type)))
# M_ids <- unique(author_vs_azimuth$author_cell_type[M_idx])
# data.table::fwrite(as.data.frame(M_ids), file = "../azimuth/Mp_author_celltypes.csv", row.names = FALSE)
# 
# # ** Monocytes ** 
# Mono <- c("Classical monocytes", "CD16- Monocyte", "Monocyte")
# Mono_idx <- unlist(sapply(Mono, function(X) grep(X, author_vs_azimuth$author_cell_type)))
# Mono_ids <- unique(author_vs_azimuth$author_cell_type[Mono_idx])
# data.table::fwrite(as.data.frame(Mono_ids), file = "../azimuth/Mo_author_celltypes.csv", row.names = FALSE)
# 
# # ** other ** 
# other <- c("CNS", "glia", "Microglia", "MDSC", "LC", "Langerhan", "microglia", "Mast cells", "MAST", "Erythroblast", "Er", "Eryth",
#            "Erythrocyte", "Neutrophil", "Basophil", "Eosinophil", "Mast", "Osteoblast", "Astrocyte",
#            "Cholangiocyte", "Acinar", "smooth", "Smooth muscle cells", "Pericyte", "muscle", "Stellate", "SMC", "PVL",
#            "stem", "GMP", "HSC", "Hepatocyte", "Keratinocyte", "Squamous", "Progenitor", "Neurons", "Normal Epithelial", "Oligo",
#            "Nueroendocrine", "Other", "Myeloid", "Myeloid cells", "Myoepithelial", "Mouse 3T3 spike-in", "MitoHigh Myeloid", 
#            "Misc/Undetermined", "Mesothelial cells", "Mesenchymal", "MSC", "Endocrine", "Doublets", "Ductal", "Epithelial",
#            "Ciliated", "Club", "CLP", "Barretts", "Unassigned", "Undetermined", "Unknown", "Chondrocyte", "other")
# 
# other_idx <- unlist(sapply(other, function(X) grep(X, author_vs_azimuth$author_cell_type)))
# other_ids <- unique(author_vs_azimuth$author_cell_type[other_idx])
# data.table::fwrite(as.data.frame(other_ids), file = "../azimuth/other_author_celltypes.csv", row.names = FALSE)
# 
# # ** Malignant **
# data.table::fwrite(as.data.frame(unique(bcmeta_annot_M$cell_type)), 
#                    file = "../azimuth/Malignant_author_celltypes.csv", row.names = FALSE)

### Azimuth ###

# # ** B cells **
# B <- c("B cell", "B")
# B_idx <- unlist(sapply(B, function(X) grep(X, author_vs_azimuth$predicted_cell_type)))
# B_ids <- unique(author_vs_azimuth$predicted_cell_type[B_idx])
# data.table::fwrite(as.data.frame(B_ids), file = "../azimuth/B_azimuth_celltypes.csv", row.names = FALSE)
# 
# # ** T cells **
# Tcell <- c("T", "T cell", "CD4", "CD8")
# T_idx <- unlist(sapply(Tcell, function(X) grep(X, author_vs_azimuth$predicted_cell_type)))
# T_ids <- unique(author_vs_azimuth$predicted_cell_type[T_idx])
# T_ids <- T_ids[-grep("IT", T_ids)]
# data.table::fwrite(as.data.frame(T_ids), file = "../azimuth/Tcell_azimuth_celltypes.csv", row.names = FALSE)
# 
# # ** NK cells **
# NK <- c("NK", "Natural")
# NK_idx <- unlist(sapply(NK, function(X) grep(X, author_vs_azimuth$predicted_cell_type)))
# NK_ids <- unique(author_vs_azimuth$predicted_cell_type[NK_idx])
# data.table::fwrite(as.data.frame(NK_ids), file = "../azimuth/NK_azimuth_celltypes.csv", row.names = FALSE)
# 
# # ** Plasma cells ** (Plasmablast considered separately)
# Plasma <- c("Plasma", "Plasmablast", "Plasmocyte", "plasmocyte")
# Plasma_idx <- unlist(sapply(Plasma, function(X) grep(X, author_vs_azimuth$predicted_cell_type)))
# Plasma_ids <- unique(author_vs_azimuth$predicted_cell_type[Plasma_idx])
# data.table::fwrite(as.data.frame(Plasma_ids), file = "../azimuth/Plasma_azimuth_celltypes.csv", row.names = FALSE)
# 
# # ** Endothelial cells ** 
# Endo <- c("Endo", "EC", "EPC", "endothelial")
# Endo_idx <- unlist(sapply(Endo, function(X) grep(X, author_vs_azimuth$predicted_cell_type)))
# Endo_ids <- unique(author_vs_azimuth$predicted_cell_type[Endo_idx])
# data.table::fwrite(as.data.frame(Endo_ids), file = "../azimuth/Endothelial_azimuth_celltypes.csv", row.names = FALSE)
# 
# # ** CAFs ** 
# CAF <- c("stromal", "fibro", "CAF", "Fibroblast", "fibroblast", "FB", "Fibro")
# CAF_idx <- unlist(sapply(CAF, function(X) grep(X, author_vs_azimuth$predicted_cell_type)))
# CAF_ids <- unique(author_vs_azimuth$predicted_cell_type[CAF_idx])
# data.table::fwrite(as.data.frame(CAF_ids), file = "../azimuth/CAFs_azimuth_celltypes.csv", row.names = FALSE)
# 
# # ** DC ** 
# DC <- c("Dendritic", "DC", "dendritic")
# DC_idx <- unlist(sapply(DC, function(X) grep(X, author_vs_azimuth$predicted_cell_type)))
# DC_ids <- unique(author_vs_azimuth$predicted_cell_type[DC_idx])
# data.table::fwrite(as.data.frame(DC_ids), file = "../azimuth/DC_azimuth_celltypes.csv", row.names = FALSE)
# 
# # ** Macrophages ** 
# M <- c("Mφ", "Macrophage", "macrophage", "mac", "TAM", "Macro", "Mac")
# M_idx <- unlist(sapply(M, function(X) grep(X, author_vs_azimuth$predicted_cell_type)))
# M_ids <- unique(author_vs_azimuth$predicted_cell_type[M_idx])
# data.table::fwrite(as.data.frame(M_ids), file = "../azimuth/Mp_azimuth_celltypes.csv", row.names = FALSE)
# 
# # ** Monocytes ** 
# Mono <- c("Classical monocytes", "CD16- Monocyte", "Monocyte", "Mono", "monocyte")
# Mono_idx <- unlist(sapply(Mono, function(X) grep(X, author_vs_azimuth$predicted_cell_type)))
# Mono_ids <- unique(author_vs_azimuth$predicted_cell_type[Mono_idx])
# data.table::fwrite(as.data.frame(Mono_ids), file = "../azimuth/Mo_azimuth_celltypes.csv", row.names = FALSE)
# 
# # ** other ** (double-check annotations, automatically not capturing everything)
# other_ids <- setdiff(author_vs_azimuth$predicted_cell_type, 
#                      c(T_ids, NK_ids, M_ids, Mono_ids, DC_ids, Endo_ids, CAF_ids, B_ids, Plasma_ids))
# data.table::fwrite(as.data.frame(other_ids), file = "../azimuth/other_azimuth_celltypes.csv", row.names = FALSE)

# ...................................... #
## Cell-type harmonization: 2º Set up final objects with both author's and azimuth's annotations
# ...................................... #

## Read harmonized cell types
author_harmonized_annots <- data.table::fread(file = "../nM_author-based_harmonization.csv")
azimuth_harmonized_annots <- data.table::fread(file = "../nM_azimuth-based_harmonization.csv")

## check all annotations are included in curated table
setdiff(unique(mapping_df$predicted_cell_type), azimuth_harmonized_annots$Name) 

## First, for confident annotations (TRUE) ##
## Add azimuth annotation (if > 1 confident annotations, keep the one with higher prediction score)

mapping_df_true <- mapping_df %>%
  dplyr::filter(!predicted_cell_type %in% c("Cycling myeloid", "immune", "Stromal") & 
                  azimuth_annotation == TRUE) %>%
  dplyr::mutate(
    predicted_cell_type_main = azimuth_harmonized_annots$cell_type_main[match(predicted_cell_type, azimuth_harmonized_annots$Name)],
    predicted_cell_type_fine = azimuth_harmonized_annots$cell_type_fine[match(predicted_cell_type, azimuth_harmonized_annots$Name)]
    ) %>%
  dplyr::group_by(cell, sample, tumor_site, author_cell_type, azimuth_annotation) %>%
  dplyr::summarise(reference = reference[which.max(prediction_score)],
                   mapping_score = mapping_score[which.max(prediction_score)],
                   azimuth_cell_type_main = predicted_cell_type_main[which.max(prediction_score)],
                   azimuth_cell_type_fine = predicted_cell_type_fine[which.max(prediction_score)],
                   prediction_score = prediction_score[which.max(prediction_score)])

## Second, for non-confident annotations
## Keep annotation based on the highest mapping score)
mapping_df_false <- mapping_df %>%
  dplyr::filter(!cell %in% unique(mapping_df_true$cell) &
                  !predicted_cell_type %in% c("Cycling myeloid", "immune", "Stromal")) %>%
  dplyr::mutate(
    predicted_cell_type_main = azimuth_harmonized_annots$cell_type_main[match(predicted_cell_type, azimuth_harmonized_annots$Name)],
    predicted_cell_type_fine = azimuth_harmonized_annots$cell_type_fine[match(predicted_cell_type, azimuth_harmonized_annots$Name)]
    ) %>%
  dplyr::group_by(cell, sample, tumor_site, author_cell_type, azimuth_annotation) %>%
  dplyr::summarise(reference = reference[which.max(mapping_score)],
                   mapping_score = mapping_score[which.max(mapping_score)],
                   azimuth_cell_type_main = predicted_cell_type_main[which.max(mapping_score)],
                   azimuth_cell_type_fine = predicted_cell_type_fine[which.max(mapping_score)],
                   prediction_score = prediction_score[which.max(mapping_score)])

## Put together
azimuth_results <- rbind(mapping_df_true, mapping_df_false)
azimuth_results <- azimuth_results %>% 
  dplyr::ungroup() %>%
  dplyr::select(!c(tumor_site, author_cell_type))

## Update metadata object with authors' manually curated annotations (NAs increased from original by adding ambiguous annotations)
harmonized_metadata_TME <- metadata_TME_v2 %>% 
  dplyr::mutate(cell_type_main = 
                  ifelse(cell_type %in% author_harmonized_annots$Name, 
                         author_harmonized_annots$cell_type_main[match(cell_type, 
                                                                  author_harmonized_annots$Name)], NA),
                cell_type_fine = 
                  ifelse(cell_type %in% author_harmonized_annots$Name, 
                         author_harmonized_annots$cell_type_fine[match(cell_type, 
                                                                  author_harmonized_annots$Name)], NA)
  )

## Update metadata object with azimuth's manually curated annotations
harmonized_metadata_TME <- inner_join(harmonized_metadata_TME, 
                                      azimuth_results,
                                      by = c("cell", "sample"))

## Add harmonized metadata to Seurat v5 object ##
## Harmonize malignant cells annotations
metadata <- merged_obj@meta.data
harmonized_metadata_M <- metadata %>% 
  #dplyr::rename("author_cell_type" = "cell_type") %>%
  dplyr::filter(malignancy == TRUE) %>%
  dplyr::mutate(cell_type_main = "Malignant",
                cell_type_fine = NA,
                azimuth_annotation = NA,
                reference = NA, 
                mapping_score = NA, 
                azimuth_cell_type_main = NA, 
                azimuth_cell_type_fine = NA, 
                prediction_score = NA)

## Put malignant and non-malignant annotations together
harmonized_metadata <- rbind(harmonized_metadata_M, harmonized_metadata_TME)

## save object
saveRDS(harmonized_metadata, file = "../azimuth/harmonized_metadata.rds")

## NA's in author's and azimuth's annotations
table(is.na(harmonized_metadata_TME$cell_type_main)) # author --> 76171 NAs
table(is.na(harmonized_metadata_TME$azimuth_cell_type_main)) # azimuth --> 0 NAs

## If we don't have confident azimuth annotations 
## which ones are missing in terms of author's? --> there would be 18448 remaining NA's
table(is.na(harmonized_metadata_TME$cell_type_main), harmonized_metadata_TME$azimuth_annotation)
ids_NAs <- which(is.na(harmonized_metadata_TME$cell_type_main) & !harmonized_metadata_TME$azimuth_annotation)
tmp_NAs <- harmonized_metadata_TME[ids_NAs,]
table(tmp_NAs$study)

## Re-annotate aml_audrey_lasry with curated author annotations ----------------

## subset aml_audrey_lasry
tmp_bcmeta <- harmonized_metadata_TME %>%
  tibble::rownames_to_column(var = "index") %>%
  filter(study == "aml_audrey_lasry") %>%
  mutate(cell_author = sapply(strsplit(cell, "_3"), head, 1),
         index = as.numeric(index))

## Add author's aml_audrey_lasry 
author_annotations_audreay_lasry <- data.table::fread(file = "../metadata_clustering_w_header_upd.csv")

## All cells present? YES
tmp_authors <- author_annotations_audreay_lasry %>% 
  filter(NAME %in% tmp_bcmeta$cell_author)

## same order as tmp_bcmeta
tmp_authors_df <- tmp_authors[match(tmp_bcmeta$cell_author, tmp_authors$NAME), ]
all.equal(tmp_authors_df$NAME, tmp_bcmeta$cell_author)

data.table::fwrite(as.data.frame(unique(tmp_authors$Broad_cell_identity)), 
                   file = "../azimuth/aml_audrey_lasry_author_celltypes.csv", row.names = FALSE)

tmp2 <- unique(tmp_authors$Broad_cell_identity)
all(tmp2 %in% author_harmonized_annots$Name)

## Add Broad_cell_identity to harmonized TME
harmonized_metadata_TME[harmonized_metadata_TME$study == "aml_audrey_lasry", "cell_type"] <- tmp_authors_df$Broad_cell_identity

harmonized_metadata_TME <- harmonized_metadata_TME %>% 
  dplyr::mutate(cell_type_main = 
                  ifelse(study == "aml_audrey_lasry",
                         author_harmonized_annots$cell_type_main[match(cell_type, author_harmonized_annots$Name)],
                         cell_type_main),
                 cell_type_fine = 
                  ifelse(study == "aml_audrey_lasry",
                         author_harmonized_annots$cell_type_fine[match(cell_type, author_harmonized_annots$Name)],
                         cell_type_fine)
  )

## Re-annotated all missing cells in aml_audrey_lasry: DONE
table(is.na(harmonized_metadata_TME$cell_type_main)) # author --> 62615 NAs
## If we don't have confident azimuth annotations, which ones are missing in terms of author's? --> 18251 NAs
table(is.na(harmonized_metadata_TME$cell_type_main), harmonized_metadata_TME$azimuth_annotation)  
ids_NAs <- which(is.na(harmonized_metadata_TME$cell_type_main) & !harmonized_metadata_TME$azimuth_annotation)
tmp_NAs <- harmonized_metadata_TME[ids_NAs,]
table(tmp_NAs$study)

## Re-annotate pdac_shu_zhang, using singleR -----------------------------------
## azimuth pancreas reference doesn't perform that good
## Pablo tested another reference dataset which wasn't accurate neither, and 
## this one was also included among the pancreas datasets used by azimuth)

## read annotations from singleR
pdac_shu_zhang_annot <- readRDS("../../data/pdac_shu_zhang_annotation_with_singleR.rds")
pdac_shu_zhang_labels_df <- pdac_shu_zhang_annot@listData %>% as.data.frame()
pdac_shu_zhang_labels_df$cell <- rownames(pdac_shu_zhang_annot)
pdac_shu_zhang_labels_df$cell <- sapply(strsplit(pdac_shu_zhang_labels_df$cell, "_"), tail ,1)

## subset pdac_shu_zhang
tmp_pdac <- harmonized_metadata_TME %>% 
  tibble::rownames_to_column(var = "index") %>%
  filter(study == "pdac_shu_zhang") %>%
  mutate(index = as.numeric(index),
         cell_author = sapply(strsplit(cell, "_"), head, 1))

## same order as tmp_bcmeta
pdac_shu_zhang_labels_df <- pdac_shu_zhang_labels_df[match(tmp_pdac$cell_author, pdac_shu_zhang_labels_df$cell), ]
## check
all.equal(pdac_shu_zhang_labels_df$cell, tmp_pdac$cell_author)

# ** pdac_shu_zhang **
data.table::fwrite(as.data.frame(unique(pdac_shu_zhang_labels_df$labels)), 
                   file = "../azimuth/pdac_shu_zhang_author_celltypes.csv", row.names = FALSE)

## read harmonized labels
singleR_harmonized_annots <- data.table::fread(file = "../nM_singleR-based_harmonization.csv")
## check
all(unique(pdac_shu_zhang_labels_df$labels) %in% singleR_harmonized_annots$Name)

## Add label to harmonized TME
harmonized_metadata_TME[harmonized_metadata_TME$study == "pdac_shu_zhang", "cell_type"] <- pdac_shu_zhang_labels_df$labels

harmonized_metadata_TME <- harmonized_metadata_TME %>% 
  dplyr::mutate(cell_type_main = 
                  ifelse(study == "pdac_shu_zhang",
                         singleR_harmonized_annots$cell_type_main[match(cell_type, singleR_harmonized_annots$Name)],
                         cell_type_main),
                 cell_type_fine = 
                  ifelse(study == "pdac_shu_zhang",
                         singleR_harmonized_annots$cell_type_fine[match(cell_type, singleR_harmonized_annots$Name)],
                         cell_type_fine)
  )

# Now 13914 cells would not have annotation
table(is.na(harmonized_metadata_TME$cell_type_main), harmonized_metadata_TME$azimuth_annotation)
ids_NAs <- which(is.na(harmonized_metadata_TME$cell_type_main) & !harmonized_metadata_TME$azimuth_annotation)
tmp_NAs <- harmonized_metadata_TME[ids_NAs,]
table(tmp_NAs$study)

## Re-annotate skcm_chao_zhang, using singleR ----------------------------------

## read annotations from singleR
skcm_chao_zhang_annot <- readRDS("../../data/skcm_chao_zhang_annotation_with_singleR.rds")
skcm_chao_zhang_labels_df <- skcm_chao_zhang_annot@listData %>% as.data.frame()
skcm_chao_zhang_labels_df$cell <- rownames(skcm_chao_zhang_annot)
skcm_chao_zhang_labels_df$cell <- sapply(strsplit(skcm_chao_zhang_labels_df$cell, "_"), tail ,1)

## subset skcm_chao_zhang
tmp_skcm <- harmonized_metadata_TME %>% 
  tibble::rownames_to_column(var = "index") %>%
  filter(study == "skcm_chao_zhang") %>%
  mutate(index = as.numeric(index),
         cell_author = sapply(strsplit(cell, "_"), head, 1))

## same order as tmp_bcmeta
skcm_chao_zhang_labels_df <- skcm_chao_zhang_labels_df[match(tmp_skcm$cell_author, skcm_chao_zhang_labels_df$cell), ]

## check
all.equal(skcm_chao_zhang_labels_df$cell, tmp_skcm$cell_author)

## harmonize cell types
singleR_harmonized_annots <- data.table::fread(file = "../nM_singleR-based_harmonization.csv")

## check
all(unique(skcm_chao_zhang_labels_df$labels) %in% singleR_harmonized_annots$Name)

## Add label to harmonized TME
harmonized_metadata_TME[harmonized_metadata_TME$study == "skcm_chao_zhang", "cell_type"] <- skcm_chao_zhang_labels_df$labels

harmonized_metadata_TME <- harmonized_metadata_TME %>% 
  dplyr::mutate(cell_type_main = 
                  ifelse(study == "skcm_chao_zhang",
                         singleR_harmonized_annots$cell_type_main[match(cell_type, singleR_harmonized_annots$Name)],
                         cell_type_main),
                 cell_type_fine = 
                  ifelse(study == "skcm_chao_zhang",
                         singleR_harmonized_annots$cell_type_fine[match(cell_type, singleR_harmonized_annots$Name)],
                         cell_type_fine)
  )

# Now 3936 cells would not have annotation
table(is.na(harmonized_metadata_TME$cell_type_main), harmonized_metadata_TME$azimuth_annotation)
ids_NAs <- which(is.na(harmonized_metadata_TME$cell_type_main) & !harmonized_metadata_TME$azimuth_annotation)
tmp_NAs <- harmonized_metadata_TME[ids_NAs,]
table(tmp_NAs$study)

## Re-annotate luad_philip_bisschof, luad_kim_nayoung --------------------------
## Decrease mapping threshold for lung reference from 0.7 to 0.6

tmp <- harmonized_metadata_TME %>% 
  filter(study %in% c("luad_philip_bisschof", "luad_kim_nayoung"))

harmonized_metadata_TME <- harmonized_metadata_TME %>%
    dplyr::mutate(azimuth_annotation = ifelse(
      study %in% c("luad_philip_bisschof", "luad_kim_nayoung") & mapping_score >= 0.6, 
      TRUE, azimuth_annotation))

# Now 2093 cells would not have annotation
table(is.na(harmonized_metadata_TME$cell_type_main), harmonized_metadata_TME$azimuth_annotation)
ids_NAs <- which(is.na(harmonized_metadata_TME$cell_type_main) & !harmonized_metadata_TME$azimuth_annotation)
tmp_NAs <- harmonized_metadata_TME[ids_NAs,]
table(tmp_NAs$study)

## For ambiguous author-based annotations, use azimuth ------------------------

celltypes_toredo <- c(
  "NK_T_cells_BUC1", "NK_T_cells_BUC2", "NK_T_cells_BUC3","NK_T_cells_BUC4","NK_T_cells_BUC5", 
  "NK_T_cells_BUC6","NK_T_cells_BUC7","NK_T_cells_BUC8","NK_T_cells_NBM1","NK_T_cells_NBM2",
  "NK_T_cells_NBM3", "T/NK cells", "T and NK", "T_cell", "T_cells", "T-cell", "T-cells", "T-cells Cycling",
  "4-Lymphocytes", "2-T cell", "30-T cell", "Memory T-Helper", "MitoHigh T-Helper", "T.cell", "Tcell_Prolif",
  "Tprolif", "TCells",
  "CNS cells", "CD163+CD14+ DCs", "Mouse 3T3 spike-in", "Other", "other", "MitoHigh Myeloid",
  "Myeloid", "Myeloid cells", 
  "Unknown", "Undetermined", "Unassigned", "Misc/Undetermined", "Doublets", "Low-quality cells",
  "Granulocyte", "Stromal", "Stromal cells", "Endocrine")

harmonized_metadata_TME <- harmonized_metadata_TME %>%
   dplyr::mutate(cell_type_main = ifelse(cell_type %in% celltypes_toredo, azimuth_cell_type_main, cell_type_main),
                 cell_type_fine = ifelse(cell_type %in% celltypes_toredo, azimuth_cell_type_fine, cell_type_fine))

## evaluate how many non-confident annotatioens are there?
tmp <- harmonized_metadata_TME %>% filter(cell_type %in% celltypes_toredo)
table(tmp$azimuth_annotation)
## 41178 FALSE vs 177172 TRUE (qualitative agreement though)

## Stick to author-based annotation, otherwise use azimuth's -------------------

harmonized_metadata_TME <- harmonized_metadata_TME %>%
  dplyr::mutate(cell_type_main = ifelse(is.na(cell_type_main) & azimuth_annotation == TRUE, 
                                        azimuth_cell_type_main, cell_type_main)) 

harmonized_metadata_TME <- harmonized_metadata_TME %>%
  dplyr::mutate(cell_type_fine = ifelse(is.na(cell_type_fine) & azimuth_annotation == TRUE, 
                                        azimuth_cell_type_fine, cell_type_fine))

table(is.na(harmonized_metadata_TME$cell_type_main)) ## Left with 2093 NAs
table(is.na(harmonized_metadata_TME$cell_type_fine)) ## Left with 2093 NAs
tmp <- harmonized_metadata_TME %>% filter(is.na(cell_type_main))


## Modify granulocyte cell_type_main -------------------------------------------
## Use instead neutrophil, basophils and eosinophil

harmonized_metadata_TME <- harmonized_metadata_TME %>%
  dplyr::mutate(cell_type_main = ifelse(cell_type_main %in% c("Neutrophil", "Basophil"),
                                        "Granulocyte", cell_type_main)) 

## Put malignant and non-malignant annotations together
harmonized_metadata <- rbind(harmonized_metadata_M, harmonized_metadata_TME)

## save object
saveRDS(harmonized_metadata_TME,
        file = "../../results/harmonized_metadata_only_nonmalignant.rds")
saveRDS(harmonized_metadata,
        file = "../../results/harmonized_metadata_malignant_and_nonmalignant.rds")
data.table::fwrite(harmonized_metadata_TME, 
                   file = "../../results/harmonized_metadata_only_nonmalignant.csv", sep = ",")
data.table::fwrite(harmonized_metadata, 
                   file = "../../results/harmonized_metadata_malignant_and_nonmalignant.csv", sep = ",")

# Now 2025 cells would not have annotation
table(is.na(harmonized_metadata_TME$cell_type_main))
ids_NAs <- which(is.na(harmonized_metadata_TME$cell_type_main))
tmp_NAs <- harmonized_metadata_TME[ids_NAs,]
table(tmp_NAs$study)

TME_annotated_df <- harmonized_metadata_TME %>% select(cell, cell_type_main, cell_type_fine)
data.table::fwrite(TME_annotated_df, 
                   file = "../../results/harmonized_bcmeta_nonmalignant_annotation.csv", sep = ",")

```

> Description plots

- Barplot (tumor types vs cell types (main, fine))
- Table (tumor types vs % cell types)

```{r}

###################################################################
## Barplot, cohort cellular description --> Top tumor types ##
###################################################################

## Use top 10 freq cancers 
freq_tumors <- c("BRCA", "LUAD", "ESCA", "GBM", "SKCM", "MM", "PAAD", "LAML", "COAD", "LUSC")

# tcca palette
tumor_type_palette <- c("GBM" = "#CA9A8C",
                        "LAML" = "#FF9A4A",
                        "BRCA" = "#FF8FAB",
                        "SKCM" = "#5E2D2C",
                        "LUSC" = "#A2F0EF",
                        "ESCA" = "#5780FE", 
                        "BLCA" = "#58B368",
                        "MM" = "#B6F884",
                        "PAAD" = "#B47EB3",
                        "COAD" = "#005D95",
                        "LUAD" = "#BBB9B7")

cell_type_palette <- c("goldenrod",
                       "#a93fc5",
                       "#5566ef",
                       "#3c8200",
                       "#c2b700",
                       "#d47bff",
                       "#01db7d",
                       "#e20644",
                       "#34dfab",
                       "#a90044",
                       "#7a92ff",
                       "salmon",
                       "#a11e14",
                       "#ff945c",
                       "#c69100",
                       "#8eb1ff",
                       "#f9632c",
                       "#008f62",
                       "#f59dff",
                       "#255e28",
                       "#c7cd4e",
                       "#857cb0",
                       "#eac250",
                       "#704274",
                       "#ba7e00",
                       "#a4615c",
                       "#f6bc6a",
                       "#8e3621",
                       "#f6b79b",
                       "#c55f00",
                       "#74490b",
                       "#aac915")

cell_type_palette <- c("#d82500",
                       "#01d582",
                       "#006cd1",
                       "#935ee7",
                       "#009f83",
                       "#b3d266",
                       "#d20086",
                       "#7dda84",
                       "#d4004a",
                       "#f15b26",
                       "#5fcaff",
                       "#b66c00",
                       "#008835",
                       "#cb4dd0",
                       "#df8f00",
                       "#7591ff",
                       "#f97d1b",
                       "#394f94",
                       "#f3be59",
                       "#8e2c68",
                       "#3d6700",
                       "#d7bafd",
                       "#b20010",
                       "#a77096",
                       "#726600",
                       "#ffabba",
                       "#993700",
                       "#acd199",
                       "#9c2432",
                       "#f7b994",
                       "#754823",
                       "#ac7f59")

names(cell_type_palette) <- c("Malignant", "Erythrocyte", "T", "Mono/Macro", "B", "NK", "Stem", "Plasmablast", "Plasma", "Megakaryocyte", "DC", "Endothelial", "Stromal", "Mast", "Endocrine", "Granulocyte", "Keratinocyte", "Osteoblast", "Epithelial", "Glial", "Exocrine", "Platelet", "Neuron", "Adipocyte", "Hepatocyte", "Cholangiocyte", "Neuroendocrine", "NKT", "Chondrocyte", "Erythroblast", "Myelocyte", "Neuroepithelial")

names(cell_type_palette) <- na.omit(unique(harmonized_metadata$cell_type_main))

## Get data 
description_tmp <- harmonized_metadata %>% 
  dplyr::filter(!is.na(cell_type_main)) %>%
  dplyr::select(cell, cell_type_main, cell_type_fine, refined_tumor_site, tumor_type) %>%
  dplyr::group_by(tumor_type, cell_type_main) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::mutate(count = seq(1, length(cell_type_main)),
                frac = n / sum(n))

description_count <- description_tmp %>%
  filter(tumor_type %in% freq_tumors) %>%
  group_by(tumor_type) %>%
  dplyr::summarise(n=n())

description_barplot <- description_tmp %>% 
  filter(tumor_type %in% freq_tumors) %>%
  ggplot(aes(x = tumor_type, y = frac, fill = cell_type_main)) +
  geom_bar(stat = "identity", position="fill") +
  scale_fill_manual(values = cell_type_palette) +
  geom_text(data = description_count, aes(x = tumor_type, y = 1.05, label = n),
            size=5, colour="black", inherit.aes=FALSE) +
  theme_classic() +
  theme(axis.title.y = element_text(size = 16, family = "Helvetica", color = "black"),
        axis.title.x = element_blank(),
        axis.text = element_text(size = 14, family = "Helvetica", color = "black"),
        legend.text = element_text(size = 12, family = "Helvetica", color = "black")) + 
  guides(fill = guide_legend("Cell type")) +
  labs(x = "Cancer type", y = "Cell type fractions")
        
###############################################################
## Table, cohort cellular description --> all tumor types ##
###############################################################

description_table <- description_tmp %>% 
  dplyr::mutate(frac = round(frac,3),
                cell_type_main = paste0("% ", cell_type_main)) %>%
  tidyr::pivot_wider(id_cols = tumor_type, names_from = cell_type_main, values_from = frac)

data.table::fwrite(x = description_table, 
                   file = "../../results/summary_table_cancertype_celltypeproportions.csv",
                   sep = ",", row.names = FALSE)

```

> TME archetype characterization

```{r}

## metadata bc-meta cohort
harmonized_metadata <- readRDS(file = "../../results/harmonized_metadata_malignant_and_nonmalignant.rds")

## Exclude liquid tumors
TME_annot <- harmonized_metadata %>%
  dplyr::select(cell, sample, tumor_type, cell_type_main, cell_type_fine) %>%
  dplyr::filter(!tumor_type %in% c("LAML", "ALL", "MM", "CLL")) %>%
  dplyr::mutate(cell_type_fine = ifelse(
    cell_type_main == "Malignant", "Malignant", cell_type_fine)
  )

# Sample-specific cell-type quantification -------------------------------------
 
## 1. Get fractions for each sample

## Main cell-type label
TME_annot_main <- TME_annot %>% 
  dplyr::filter(!is.na(cell_type_main)) %>%
  dplyr::select(cell, sample, tumor_type, cell_type_main) %>%
  dplyr::group_by(sample, tumor_type, cell_type_main) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::mutate(count = seq(1, length(cell_type_main)),
                frac = n / sum(n)) %>%
  dplyr::mutate(three_feature_TME = NA)

## Fine cell-type label
TME_annot_fine <- TME_annot %>% 
  dplyr::filter(!is.na(cell_type_main)) %>%
  dplyr::select(cell, sample, tumor_type, cell_type_fine) %>%
  dplyr::group_by(sample, tumor_type, cell_type_fine) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::mutate(count = seq(1, length(cell_type_fine)),
                frac = n / sum(n)) %>%
  dplyr::mutate(eight_feature_TME = NA)

# --------------------- #
### Stroma annotation ###
# --------------------- #

stromal_idx <- grep("Stromal", TME_annot_main$cell_type_main)
TME_annot_main$three_feature_TME[as.numeric(stromal_idx)] <- "Stromal"

# --------------------- #
### Lymphoid annotation ###
# --------------------- #

Lymphoid <- c("B", "T", "NK", "Plasma", "Plasmablast", "MAIT", "NKT")
lymphoid_idx <- do.call(c, sapply(Lymphoid, function(X) grep(X, TME_annot_main$cell_type_main)))
TME_annot_main$three_feature_TME[as.numeric(lymphoid_idx)] <- "Lymphoid"

# --------------------- #
### Myeloid annotation ###
# --------------------- #

Myeloid  <- c("Mono/Macro", "DC", "Mast", "Erythroblast", 
              "Platelet", "Erythrocyte", "Granulocyte", "Megakaryocyte")
myeloid_idx <- do.call(c, sapply(Myeloid, function(X) grep(X, TME_annot_main$cell_type_main)))
TME_annot_main$three_feature_TME[as.numeric(myeloid_idx)] <- "Myeloid"

## cellines are removed here
TME_annot_main <- TME_annot_main %>% 
  filter(!is.na(three_feature_TME)) %>%
  select(!c(n, count, cell_type_main))

# --------------------- #
### Treg annotation ###
# --------------------- #

treg <- c("Treg")
treg_idx <- do.call(c, lapply(treg, function(X) grep(X, TME_annot_fine$cell_type_fine)))
TME_annot_fine$eight_feature_TME[as.numeric(treg_idx)] <- "Treg"

# --------------------- #
### CD4 T annotation ###
# --------------------- #

cd4t <- c("CD4", "Th")
cd4t_idx <- do.call(c, sapply(cd4t, function(X) grep(X, TME_annot_fine$cell_type_fine)))
TME_annot_fine$eight_feature_TME[as.numeric(cd4t_idx)] <- "CD4T"

# --------------------- #
### CD8 T annotation ###
# --------------------- #

cd8t <- c("CD8")
cd8t_idx <- do.call(c, lapply(cd8t, function(X) grep(X, TME_annot_fine$cell_type_fine)))
TME_annot_fine$eight_feature_TME[as.numeric(cd8t_idx)] <- "CD8T"

# --------------------- #
### Mo annotation ###
# --------------------- #

mo <- c("Mo")
mo_idx <- do.call(c, lapply(mo, function(X) grep(X, TME_annot_fine$cell_type_fine)))
TME_annot_fine$eight_feature_TME[as.numeric(mo_idx)] <- "Mo"

# --------------------- #
### Mp annotation ###
# --------------------- #

mp <- c("Mp")
mp_idx <- do.call(c, lapply(mp, function(X) grep(X, TME_annot_fine$cell_type_fine)))
TME_annot_fine$eight_feature_TME[as.numeric(mp_idx)] <- "Mp"

# --------------------- #
### cDC1 annotation ###
# --------------------- #

dc1 <- c("cDC1")
dc1_idx <- do.call(c, lapply(dc1, function(X) grep(X, TME_annot_fine$cell_type_fine)))
TME_annot_fine$eight_feature_TME[as.numeric(dc1_idx)] <- "cDC1"

# --------------------- #
### cDC2 annotation ###
# --------------------- #

dc2 <- c("cDC2")
dc2_idx <- do.call(c, lapply(dc2, function(X) grep(X, TME_annot_fine$cell_type_fine)))
TME_annot_fine$eight_feature_TME[as.numeric(dc2_idx)] <- "cDC2"

# --------------------- #
### endothelial annotation ###
# --------------------- #

endo <- c("Endothelial")
endo_idx <- do.call(c, lapply(endo, function(X) grep(X, TME_annot_fine$cell_type_fine)))
TME_annot_fine$eight_feature_TME[as.numeric(endo_idx)] <- "Endothelial"

# --------------------- #
### CAF annotation ###
# --------------------- #

caf <- c("CAF")
caf_idx <- do.call(c, lapply(caf, function(X) grep(X, TME_annot_fine$cell_type_fine)))
TME_annot_fine$eight_feature_TME[as.numeric(caf_idx)] <- "CAF"

TME_annot_fine <- TME_annot_fine %>% 
  filter(!is.na(eight_feature_TME)) %>%
  select(!c(n, count, cell_type_fine))

## Put together all features in a data.frame ##

mat_TME_main <- TME_annot_main %>%
  dplyr::group_by(sample, tumor_type, three_feature_TME) %>%
  dplyr::summarise(frac_feature = sum(frac)) %>% 
  dplyr::filter(!is.na(three_feature_TME)) %>%
  tidyr::pivot_wider(id_cols = "sample", 
                     names_from = three_feature_TME,
                     values_from = frac_feature)

mat_TME_fine <- TME_annot_fine %>%
  dplyr::group_by(sample, tumor_type, eight_feature_TME) %>%
  dplyr::summarise(frac_feature = sum(frac)) %>% 
  dplyr::filter(!is.na(eight_feature_TME)) %>%
  tidyr::pivot_wider(id_cols = "sample", 
                     names_from = c(eight_feature_TME),
                     values_from = frac_feature) 

mat_TME <- left_join(mat_TME_main, mat_TME_fine)
mat_TME <- mat_TME %>%
  tibble::column_to_rownames(var = "sample") %>% 
  as.data.frame()

## set cell-type NAs set to 0,
## assumption no measure = no presence rather than we don't have that information)

mat_TME[is.na(mat_TME)] <- 0
apply(mat_TME, 2, summary)
range(mat_TME)

mat_TME <- mat_TME[,c("Lymphoid", "Myeloid", "Stromal","Treg", "CD8T", "CD4T", "Mp", "Mo", "cDC1", "cDC2", "Endothelial", "CAF")]
saveRDS(mat_TME, file = "../../results/bcmetaTME_scaled_cellfrac_heatmap.rds")

```

> Inspect communities based on TME features

1. Define 10 TME features
2. Neighborhood graph based KNN using euclidean distances. 
k [3, 4, 5, …, 300]
3. Identification of clusters using the Louvain method (i.e. maximizes network modularity)
At each k, res [0.3, 0.4, 0.5, …, 2]
4. Resulting clustering evaluated using the Davies Bouldin Index (DBI) which assess the ratio of the intra-cluster distance to inter-cluster distance. 

Low DBI → Better separation between clusters, and more compact clusters

```{r}

## packages
library('ConsensusClusterPlus')
library('igraph')
library('ggridges')
library('GGally')
library('dendextend')
library('umap')
library('NbClust')
library('clv')

## Visualize feature distributions ---------------------------------------------

## TME matrix
mat_TME <- readRDS(file = "../../results/bcmetaTME_scaled_cellfrac_heatmap.rds")

## to data.frame
mat_TME_df <- mat_TME  %>%
  tibble::rownames_to_column(var = "sample") %>%
  tidyr::pivot_longer(cols = !sample, 
                      names_to = "TME_feature", 
                      values_to = "frac")

mat_TME_df$TME_feature <- factor(mat_TME_df$TME_feature,
                                 levels = colnames(mat_TME))

mat_breaks <- seq(min(mat_TME, na.rm = TRUE), max(mat_TME, na.rm = TRUE), length.out = 100)
col_fun <- viridisLite::rocket(length(mat_breaks) - 1, direction = -1)

## TME features palette
tme_palette <- c("#2d7600",
                 "#6f5be4",
                 "#01d08a",
                 "#ff78ee",
                 "#778200",
                 "#2991ff",
                 "#f57b15",
                 "#39d9ec",
                 "#d30036",
                 "#87935d",
                 "#ff8575",
                 "#946344")

names(tme_palette) <- colnames(mat_TME)

mat_TME_df %>%
  ggplot(aes(x = frac, y = TME_feature, color = TME_feature, fill = TME_feature)) +
  #geom_density_ridges() +
  geom_boxplot() +
  theme_classic() + 
  coord_flip() +
  scale_fill_manual(values = tme_palette, name = "TME feature") +
  scale_color_manual(values = tme_palette, name = "TME feature") +
  theme(axis.title = element_blank(), 
        axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = "black"),
        axis.text.y = element_text(size = 12, angle = 0, color = "black")) +
  labs(x = "Proportions") +
  guides(fill = "none", color = "none")

## Definition of TME archetypes ------------------------------------------------

## Annotation
harmonized_metadata <- readRDS(file = "../../results/harmonized_metadata_malignant_and_nonmalignant.rds")

## remove liquid tumors
TME_annot <- harmonized_metadata %>%
  dplyr::select(cell, sample, tumor_type, cell_type_main, cell_type_fine) %>%
  dplyr::filter(!tumor_type %in% c("LAML", "ALL", "MM", "CLL")) %>%
  dplyr::mutate(cell_type_fine = 
                  ifelse(cell_type_main == "Malignant", "Malignant", cell_type_fine)
  )

## sample annotations
annot <- TME_annot %>% 
  dplyr::group_by(sample, tumor_type) %>%
  dplyr::summarise() %>% 
  tibble::column_to_rownames(var = "sample") %>%
  dplyr::rename("Tumor type" = "tumor_type")

obj_TME <- CreateSeuratObject(counts = t(mat_TME), 
                              meta.data = annot,
                              project = "TME",
                              assay = "clustering",
                              min.cells = 0,
                              min.features = 0,
                              names.field = 2,
)

## set data slot = counts lost
obj_TME@assays$clustering$data <- obj_TME@assays$clustering$counts

## Look into PCA space
obj_TME <- FindVariableFeatures(obj_TME)
obj_TME <- ScaleData(obj_TME)
obj_TME <- RunPCA(obj_TME, npcs = 50)

VizDimLoadings(obj_TME, dims = 1:11, reduction = "pca")
DimHeatmap(obj_TME, dims = 1:11, balanced = TRUE)
ElbowPlot(obj_TME, ndims = 11)

saveRDS(object = obj_TME, file = "../../results/object_seurat_TME_features.rds")

## Identify optimal clustering parameters --------------------------------------
## Parameters k neighbours, res resolution

# obj <- readRDS()
# obj_TME <- lapply(3:300, function(X){
#   obj_TME <- lapply(seq(.3,2,0.1), function(Y){
# 
#     obj <- FindNeighbors(obj, dims = 1:10, reduction = "pca", k.param = X)
#     obj <- FindClusters(obj, cluster.name = "clusters", resolution = Y)
#     dbi_perf <- evaluate_clustering(data = obj_TME@reductions$pca@cell.embeddings, membership = obj$seurat_clusters)
# 
#   return(list(object_seurat = obj, dbi_perf = dbi_perf))
#   })
#   return(obj_TME)
# })

## Evaluate clustering results according to DBI --------------------------------

# evaluate_clustering <- function(data, membership){
# 
#   intraclust = c("centroid")
#   interclust = c("centroid")
# 
#   if (length(unique(membership)) > 1) {
# 
#     clust_pred <- as.integer(membership)
# 
#     # compute intercluster distances and intracluster diameters
#     cls.scatt <- cls.scatt.data(data, clust_pred, dist="euclidean")
# 
#     # compute Davies-Bouldin index
#     dbi_out <- clv.Davies.Bouldin(cls.scatt, intraclust, interclust)
# 
#     return(dbi_out)
#   }else{
#     return(NULL)
#   }
# }
       
## Load performance results ----------------------------------------------------
## Parameters k neighbours, res resolution that yielded to lowest DBI
louvain_res <- readRDS(file = "../../results/lovuain_pca_space.rds")

tmp <- lapply(1:length(3:300), function(X){
  tmp <- lapply(1:length(seq(.3,2,0.1)), function(Y){
    return(louvain_res[[X]][[Y]]$dbi_perf)
  })
  names(tmp) <- seq(.3,2,0.1)
  return(tmp)
})
names(tmp) <- 3:300

tmp2 <- unlist(tmp)

## number of neighbours, k = 7
k <- sapply(strsplit(names(which.min(tmp2)), split = ".", fixed = TRUE), head, 1)

## resolution parameter, res = 0.7
res <- sapply(strsplit(names(which.min(tmp2)), split = ".", fixed = TRUE), tail, 2)
res <- paste(res, collapse = ".")

## Neighbourhood graph (KNN) based on the euclidean distance in PCA space
obj_TME <- FindNeighbors(obj_TME, dims = 1:10, reduction = "pca", k.param = as.numeric(k))

## Definition of clusters based on Louvain algorithm, maximizing modularity
obj_TME <- FindClusters(obj_TME, cluster.name = "clusters", algorithm = 1, resolution = as.numeric(res))

color_clusters <- c("#c9593d",
                    "#60b34d",
                    "#9a56c8",
                    "#b5b342",
                    "#6e6fc7",
                    "#c58942",
                    "#cd62ba",
                    "#4bb292",
                    "#d2406c",
                    "#687a37",
                    "#6a99d3",
                    "#bb688b")

## UMAP plot
obj_TME <- RunUMAP(obj_TME, dims = 1:10, reduction = "pca", reduction.name = "umap")
DimPlot(obj_TME, cols = color_clusters)

## PCA plot
DimPlot(obj_TME, reduction = "pca", cols = color_clusters)

# find all markers of each cluster
cluster_markers <- lapply(0:11, function(X){
  markers <- FindMarkers(obj_TME, ident.1 = X)
  return(markers)
})
names(cluster_markers) <- 0:11

markers <- FindMarkers(obj_TME, ident.1 = 0, ident.2 = 1)

## Heatmap - markers
cluster_makers_df <- do.call(rbind, lapply(0:11, function(X){
  tmp <- cluster_markers[[as.character(X)]]
  tmp$cluster <- X
  tmp <- tmp %>% 
    tibble::rownames_to_column(var = "TME_feature")
  return(tmp)
}))

cluster_makers_df$TME_feature <- factor(cluster_makers_df$TME_feature, 
                                        levels = rev(c("Lymphoid", "CD8T", "CD4T", "Treg",
                                                       "Stromal", "CAF", "Endothelial", "Myeloid", "Mp", "Mo",
                                                       "cDC1","cDC2")))

lim <- min(range(cluster_makers_df$avg_log2FC))
cluster_makers_df$avg_log2FC[which(cluster_makers_df$avg_log2FC > abs(lim))] <- abs(lim)

cluster_makers_df %>%
  ggplot(aes(x=as.factor(cluster), y=TME_feature, fill = avg_log2FC)) +
  geom_tile() +
  scale_fill_gradient2(low ="blue", mid = "white", high = "red", midpoint = 0) +
  theme_classic() +
  theme(axis.title = element_blank(), axis.text = element_text(size = 14, color = "black")) +
  labs(x="Cluster", y = "TME feature", fill = "Average\nlog2\nFold-Change") +
  geom_text(aes(label=ifelse(p_val_adj > .05, "X", "")))

## Heatmap - samples

mat_breaks <- seq(min(mat_TME, na.rm = TRUE), max(mat_TME, na.rm = TRUE), length.out = 100)
mat_breaks[100] <- round(mat_breaks[100] )
col_fun <- viridisLite::rocket(length(mat_breaks) - 1, direction = -1)

heatmap <- DoHeatmap(obj_TME, 
                     slot = "data", 
                     features = rev(levels(cluster_makers_df$TME_feature)),
                     group.colors = color_clusters,
) + 
  scale_fill_gradientn(colours = col_fun, values = mat_breaks) +
  theme(axis.text.y = element_text(size = 12, color = "black"),
        legend.text = element_text(size = 10, color = "black"))

## Visualize feature distributions per cluster ---------------------------------

obj_TME_cluster_1 <- subset(obj_TME, seurat_clusters == 1)
cluster1 <- WhichCells(object = obj_TME_cluster_1)
obj_TME_cluster_1 <- SetIdent(obj_TME_cluster_1, cells = cluster1, value = 1)

RidgePlot(obj_TME_cluster_1, features = c("Myeloid", "Mp"), ncol = 4, idents = 1)

## Establish a TME archetype for each sample -----------------------------------

bcmeta_archetypes <- obj_TME@meta.data %>%
  tibble::rownames_to_column(var = "sample") %>%
  dplyr::mutate(archetype = as.numeric(seurat_clusters))

bcmeta_archetypes$archetype <- factor(bcmeta_archetypes$archetype, levels = sort(unique(bcmeta_archetypes$archetype)))

summary <- bcmeta_archetypes %>%
  dplyr::group_by(`Tumor type`, archetype) %>%
  dplyr::summarise(count = n()) %>%
  ggplot(aes(x = archetype, y=`Tumor type`, fill = count)) +
  ggplot2::geom_tile() + 
  ggplot2::scale_fill_gradient(low = "white", high = "darkgreen") +
  ggplot2::geom_text(aes(label = count), position = position_dodge2(0.2)) +
  theme_classic() +
  theme(axis.title = element_blank(), 
        axis.text = element_text(size = 16, colour = "black"))
  # scale_x_discrete(breaks = levels(bcmeta_archetypes$archetype),
  #                  limits = c("skip", 
  #                             levels(bcmeta_archetypes$archetype)[1], "skip", "skip",
  #                             levels(bcmeta_archetypes$archetype)[2], "skip", "skip", 
  #                             levels(bcmeta_archetypes$archetype)[3], "skip", "skip", 
  #                             levels(bcmeta_archetypes$archetype)[4], "skip", "skip", 
  #                             levels(bcmeta_archetypes$archetype)[5], "skip", "skip", 
  #                             levels(bcmeta_archetypes$archetype)[6], "skip", "skip", 
  #                             levels(bcmeta_archetypes$archetype)[7], "skip", "skip", 
  #                             levels(bcmeta_archetypes$archetype)[8], "skip", "skip",
  #                             levels(bcmeta_archetypes$archetype)[9], "skip", 
  #                             levels(bcmeta_archetypes$archetype)[10],
  #                             levels(bcmeta_archetypes$archetype)[11],
  #                             levels(bcmeta_archetypes$archetype)[12]
  #                             ))

summary
ggsave(filename = "../../figures/summary_samples_TME_archetypes.png", width = 14, height = 8)


ggpubr::ggarrange(heatmap, summary, heights = c(1.5, 1), ncol = 1, nrow = 2, align = "v", )
heatmap / summary

bcmeta_archetypes <- bcmeta_archetypes %>%
  dplyr::mutate(archetype = ifelse(archetype == 1, "Myeloid_centric", archetype)) %>%
  dplyr::mutate(archetype = ifelse(archetype == 2, "Immune_rich_Treg_bias", archetype)) %>%
  dplyr::mutate(archetype = ifelse(archetype == 3, "Immune_rich_cDC1_bias", archetype)) %>%
  dplyr::mutate(archetype = ifelse(archetype == 4, "Immune_desert_cDC1_bias", archetype)) %>%
  dplyr::mutate(archetype = ifelse(archetype == 5, "Immune_stromal_CAFlike_cDC1_cDC2_bias", archetype)) %>%
  dplyr::mutate(archetype = ifelse(archetype == 6, "Immune_stromal_Endolike_Treg_cDC1_cDC2_bias", archetype)) %>%
  dplyr::mutate(archetype = ifelse(archetype == 7, "Immune_stromal_cDC1_bias", archetype)) %>%
  dplyr::mutate(archetype = ifelse(archetype == 8, "Tcell_centric_Mo_cDC1_cDC2_bias", archetype)) %>%
  dplyr::mutate(archetype = ifelse(archetype == 9, "Myeloid_centric_Mp_bias", archetype)) %>%
  dplyr::mutate(archetype = ifelse(archetype == 10, "Immune_desert_Endo_Mo_cDC2_bias", archetype)) %>%
  dplyr::mutate(archetype = ifelse(archetype == 11, "Myeloid_centric_Mo_cDC1_cDC2_bias", archetype)) %>%
  dplyr::mutate(archetype = ifelse(archetype == 12, "Immune_stromal_CAFlike", archetype)) %>%
  dplyr::select(sample, archetype) %>%
  dplyr::rename("TME_archetype" = "archetype")

saveRDS(bcmeta_archetypes, file = "../../results/bcmeta_samples_TME_archetype.rds")
data.table::fwrite(bcmeta_archetypes, 
                   file = "../../results/bcmeta_samples_TME_archetype.csv",
                   row.names = FALSE,
                   quote = FALSE
                   )

```
                                    
> Investigate association between metacomunidades and TME archetipos (untreated-primary, BrM por separado)

```{r}

## Without BrMs ----------------------------------------------------------------

## TME archetypes
TME_annot <- harmonized_metadata %>%
  dplyr::group_by(sample, tumor_type, treated, sample_type) %>%
  dplyr::filter(sample %in% rownames(mat_TME) & !study %in% c("brmets_hugo_gonzalez", "brmets_jana_biermann")) %>%
  dplyr::summarise()

TME_annot <- inner_join(TME_annot, bcmeta_archetypes, by = "sample")
TME_annot$TME_archetype <- gsub("_", "\n", TME_annot$TME_archetype)
TME_annot$treated[TME_annot$treated == ""] <- NA

TME_annot <- TME_annot %>%
  dplyr::rename("Sample" = "sample",
                "CancerType" = "tumor_type",
                "SampleType" = "sample_type",
                "Treated" = "treated",
                "TMEarchetype" = "TME_archetype") 

TME_annot <- TME_annot %>%
  dplyr::ungroup() %>%
  dplyr::select(Sample, TMEarchetype)

## Therapeutic metacommunities
drug_metacommunities <- data.table::fread("../../results/bcmeta_metacommunities.csv")
drug_metacommunities <- drug_metacommunities %>%
  dplyr::select(Sample, Study, `Cancer type`, Subtype, `Primary or metastasis`, Treated, `Best metacommunity by frequency`) %>%
  dplyr::filter(Sample %in% rownames(mat_TME) & !Study %in% c("brmets_hugo_gonzalez", "brmets_jana_biermann")) %>%
  dplyr::rename("CancerType" = `Cancer type`,
                "SampleType" = `Primary or metastasis`,
                "MetaCommunity" = `Best metacommunity by frequency`)

df_drug_TME <- inner_join(drug_metacommunities, TME_annot, by = c("Sample"))

#########################
## Primary/Metastasis TME archetypes
#########################

dat_origin <- table(df_drug_TME$SampleType, df_drug_TME$TMEarchetype)

## Chi-square test
chisq.test(dat_origin, correct = FALSE)

## Expected frequencies
chisq.test(dat_origin)$expected

## Fisher's exact test
test <- fisher.test(dat_origin, workspace = 2e8)

# combine plot and statistical test with ggbarstats
library('ggstatsplot')

ggbarstats(
  df_drug_TME, SampleType, TMEarchetype,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(test$p.value, 3))
  )
)  +
  theme(axis.title = element_blank(), 
        axis.ticks.x = element_blank(), 
        axis.text.y = element_text(size = 12, colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black", hjust = 0.5))

ggsave(filename = "../../figures/PrimaryMetastasis_vs_TMEarchetype.png", width = 12, height = 8)

#########################
## Treatment vs TME archetypes
#########################

tmp <- df_drug_TME %>%
  dplyr::filter(!Treated %in% c("Unknown"))

dat_treatment <- table(tmp$Treated, tmp$TMEarchetype)

## Chi-square test
chisq.test(dat_treatment, correct = FALSE)

## Expected frequencies
chisq.test(dat_treatment)$expected

## Fisher's exact test
test <- fisher.test(dat_treatment, workspace = 2e9)

# combine plot and statistical test with ggbarstats

ggbarstats(
  tmp, Treated, TMEarchetype,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(test$p.value, 3))
  )
)  +
  theme(axis.title = element_blank(), 
        axis.ticks.x = element_blank(), 
        axis.text.y = element_text(size = 12, colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black", hjust = 0.5))

ggsave(filename = "../../figures/TreatmentStatus_vs_TMEarchetype.png", width = 12, height = 8)

#########################
## Metacommunities vs TME archetypes
#########################

dat_drug <- table(df_drug_TME$MetaCommunity, df_drug_TME$TMEarchetype)

## Chi-square test
chisq_test <- chisq.test(dat_drug, correct = FALSE)

## Expected frequencies
chisq.test(dat_drug)$expected

## Fisher's exact test --> leads to an error
test <- fisher.test(dat_drug)

# combine plot and statistical test with ggbarstats

ggbarstats(
  df_drug_TME, MetaCommunity, TMEarchetype,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Chi-square test", ", p-value = ",
    ifelse(chisq_test$p.value < 0.001, "< 0.001", round(chisq_test$p.value, 3))
  )
)  +
  theme(axis.title = element_blank(), 
        axis.ticks.x = element_blank(), 
        axis.text.y = element_text(size = 12, colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black", hjust = 0.5))

ggsave(filename = "../../figures/MetaCommunity_vs_TMEarchetype.png", width = 12, height = 8)

## With BrMs ----------------------------------------------------------------

## TME archetypes
TME_annot <- harmonized_metadata %>%
  dplyr::group_by(sample, tumor_type, treated, sample_type) %>%
  dplyr::filter(sample %in% rownames(mat_TME) & study %in% c("brmets_hugo_gonzalez", "brmets_jana_biermann")) %>%
  dplyr::summarise()

TME_annot <- inner_join(TME_annot, bcmeta_archetypes, by = "sample")
TME_annot$TME_archetype <- gsub("_", "\n", TME_annot$TME_archetype)
TME_annot$treated[TME_annot$treated == ""] <- NA

TME_annot <- TME_annot %>%
  dplyr::rename("Sample" = "sample",
                "CancerType" = "tumor_type",
                "SampleType" = "sample_type",
                "Treated" = "treated",
                "TMEarchetype" = "TME_archetype") 

TME_annot <- TME_annot %>%
  dplyr::ungroup() %>%
  dplyr::select(Sample, TMEarchetype)

## Therapeutic metacommunities
drug_metacommunities <- data.table::fread("../../results/bcmeta_metacommunities.csv")
drug_metacommunities <- drug_metacommunities %>%
  dplyr::select(Sample, Study, `Cancer type`, Subtype, `Primary or metastasis`, Treated, `Best metacommunity by frequency`) %>%
  dplyr::filter(Sample %in% rownames(mat_TME) & Study %in% c("brmets_hugo_gonzalez", "brmets_jana_biermann")) %>%
  dplyr::rename("CancerType" = `Cancer type`,
                "SampleType" = `Primary or metastasis`,
                "MetaCommunity" = `Best metacommunity by frequency`)

df_drug_TME <- inner_join(drug_metacommunities, TME_annot, by = c("Sample"))

#########################
## Treatment vs TME archetype
#########################

dat_treatment <- table(df_drug_TME$Treated, df_drug_TME$TMEarchetype)

## Chi-square test
chisq.test(dat_treatment, correct = FALSE)

## Expected frequencies
chisq.test(dat_treatment)$expected

## Fisher's exact test
test <- fisher.test(dat_treatment, workspace = 2e9)

# combine plot and statistical test with ggbarstats

ggbarstats(
  df_drug_TME, Treated, TMEarchetype,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(test$p.value, 3))
  )
)  +
  theme(axis.title = element_blank(), 
        axis.ticks.x = element_blank(), 
        axis.text.y = element_text(size = 12, colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black", hjust = 0.5))

ggsave(filename = "../../figures/TreatmentStatus_vs_TMEarchetype_BrMs.png", width = 8, height = 8)

#########################
## Metacommunities vs TME archetypes
#########################

dat_drug <- table(df_drug_TME$MetaCommunity, df_drug_TME$TMEarchetype)

## Chi-square test
chisq_test <- chisq.test(dat_drug, correct = FALSE)

## Fisher's exact test --> leads to an error
fisher_test <- fisher.test(dat_drug)

# combine plot and statistical test with ggbarstats

ggbarstats(
  df_drug_TME, MetaCommunity, TMEarchetype,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's Exact Test", ", p-value = ",
    ifelse(fisher_test$p.value < 0.001, "< 0.001", round(fisher_test$p.value, 3))
  )
)  +
  theme(axis.title = element_blank(), 
        axis.ticks.x = element_blank(), 
        axis.text.y = element_text(size = 12, colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black", hjust = 0.5))

ggsave(filename = "../../figures/MetaCommunity_vs_TMEarchetype_BrMs.png", width = 8, height = 8)

```
